---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# adjHelpR

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/gi0na/adjHelpR.svg?branch=master)](https://travis-ci.com/gi0na/adjHelpR)
<!-- badges: end -->

The goal of adjHelpR is to provide tools to generate adjacency matrices from weighted edgelists, create weighted edgelists from edgelists with repeated edges, and handle these objects.
It exploits `dplyr` and `tibble`s to deal efficiently with large edgelists, and `Matrix` sparse matrices to store adjacency matrices in sparse format.

## Installation

You can install the released version of adjHelpR from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("adjHelpR")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("gi0na/adjHelpR")
```
## Example

This is a basic example on how to generate a weighted adjacency matrix from an edge list:

```{r example}
library(adjHelpR)

el <- data.frame(from= c('a','b','b','c','d','d'),
                 to  = c('b','c','d','a','b','a'),
                 attr= c( 12, 6, 12 , 6 , 6 , 6 ))
(adj <- get_adjacency(el))

(el_unweighted <- tibble::as_tibble(do.call(rbind, 
                        apply(el, 1, function(row) matrix(rep(row[1:2], each=as.integer(row[3])), ncol = 2))), 
                        .name_repair='minimal'))
(adj <- get_adjacency(el_unweighted, multiedge = TRUE))
```

# Acknowledgements
The research and development behind `adjHelpR` is performed at the Chair of Systems Design, ETH Zürich.

# Contributors

[Giona Casiraghi](http://giona.info)


# Copyright
`adjHelpR` is licensed under the [GNU Affero General Public License](https://choosealicense.com/licenses/agpl-3.0/).

\(c\) Copyright ETH Zürich, 2019-2020
